Phenomenological Robotics

### 2011-12-21 draft.tex
Wrote on Van Gelder coupling, Freeman attractors and the relation to Merleau-Ponty.

>>>>

% 2.C. Merleau-Ponty vs. Coupling (van Gelder) vs. Stepp vs. Freeman.
% (Source: Thesis journal, pages 124 to 129, 11/03/2011).
% (Source: Thesis journal, pages 90 to 99, 18/02/2011).
% Note: Van Gelder ``coupling'' seems to be directly related to Freeman's basins
% of attraction. I don't see how Stepp would come between them. 2.D is supposed
% to be EST vs. discrete basin transitions, so it might belong down there. Maybe
% we could cite our work on EST that also includes coupling.

Van Gelder~\cite{vangelder1998} defends the \emph{dynamical hypothesis} in
cognitive science, stating that

\begin{quotation}
  \dots cognitive agents ``are'' \dots not some particular dynamical systems but
  as many systems as are needed to produce \dots cognitive performances \dots
\end{quotation}

He proposes that a cognitive system interacts with the environment through
\emph{coupling}, with input influencing the direction of change continuously and
output doing the same to the external world. Two different systems forging each
other's transformations at the same time.

Freeman~\cite{freeman1991}, after a set of experiments with rabbits, developed
an \emph{attractor theory}, where the brain can be understood as a dynamical 
system. According to this theory, the brain is composed of a set of energy peaks
and valleys (formed from one's past experiences). When the subject is in a
specific type of situation, the corresponding part of the brain landscape is
activated, and movements are caused to get the brain closer to a state of
minimum energy, called a ``basin of attraction''.

Van Gelder's idea of a cognitive system being direcly connected to the world can
be seen as an instance of Merleau-Ponty's \emph{intentional arc}, where
Freeman's results support the validity of the \emph{maximal grip} concept.
